---
title: "MobiLipid"
output:   
  md_document:
    variant: markdown_github
abstract: 'The goal of MobiLipid is to monitor the bias of collisional cross section (CCS) when using secondary ion mobility technologies (e.g. trapped ion mobility spectrometry (TIMS)). MobiLipid can be used when 
'
---

# MobiLipid

## Necessary installations

To run MobiLipid it is necessary to install [R](https://cran.r-project.org/) and [R studio](https://www.rstudio.com/products/rstudio/download/).

Afterwards, it has to be ensured that all necessary packages, which are needed to run MobiLipid, are installed. Therefore, paste the following code into the console of R studio and press enter.

```{r, include=TRUE, eval=FALSE}
# List of packages to check/install
packages <- c("htmltools" ,"rmarkdown", "data.table", "ggplot2", "DT", "webshot", "tcltk", "knitr", "ggbeeswarm")

# Check if packages are installed, if not, install them
for (package in packages) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
}
```

## Download of MobiLipid

Furthermore, it has to be ensured that the .Rmd file containing MobiLipid ("MobiLipid_CCS-bias-calculation.Rmd" and/or "MobiLipid_CCS-bias-calculation_CCS-correction.Rmd") as well as the .csv file of the ^DT^CCS~N2~ library for U^13^C labeled lipids of yeast ("U13C_DT_CCS_library.csv") is downloaded. This can be done by downloading the entire GitHup repository as a .zip file and unzip the files. 

## Running MobiLipid

Finally, MobiLipid can be run. Therefore, paste the code below into the console of R studio and press enter. The code will open 3 pop-up windows where you have to choose the following files (be aware that the windows might open behind the main window of R studio):

1. **Select the R markdown which should be used for data processing.** This needs to be a .Rmd file. "MobiLipid_CCS-bias-calculation.Rmd" can be used to calculate the CCS bias betweent measured CCS values and ^DT^CCS~N2~ library values of U^13^C labeled yeast lipids or "MobiLipid_CCS-bias-calculation_CCS-correction.Rmd" to additionally perform a CCS correction based on linear correction functions using the ^DT^CCS~N2~ library.
2. **Data import (measured data as .csv file)**. This needs to be a .csv file containing the measured data. The .csv file has to have the following headers: File, LipidClass, LipidSpecies, Adduct, Label, IMS_value, CCS (an examplary file es provided: "Example_data_import.csv")
    - **File**: File name of the measurement file
    - **LipidClass**: Lipid class (ensure to use the same nomenclature as used for the ^DT^CCS~N2~ library)
    - **LipidSpecies**: Lipid species on lipid species level (ensure to use the same nomenclature as used for the ^DT^CCS~N2~ library)
    - **Adduct**: The following adducts are possible: [M+H], [M+NH~4~], [M+Na], [M-H], and [M+HCOO]. Not all adducts can be used for each lipid class. The table below shows the possible lipid class-adduct combinations:
    ```{r, include=TRUE, eval=TRUE, echo=FALSE}
class_adduct <- data.table::data.table("Lipid class" = c("AcCa", "Cer", "Co", "DG", "HexCer", "LPC", "LPE", "PA", "PC", "PE", "PG", "PI", "PS", "SPH", "TG"), "Adduct" = c("[M+H], [M-H]", "[M+H], [M+Na], [M+HCOO]", "[M+NH4], [M+Na]", "[M+NH4], [M+Na]", "[M+H], [M+Na], [M+HCOO]", "[M+H], [M+Na], [M+HCOO]", "[M+H], [M-H]", "[M+NH4], [M+Na], [M-H]", "[M+H], [M+Na], [M+HCOO]", "[M+H], [M+Na], [M-H]", "[M+NH4], [M-H]", "[M+NH4], [M+Na], [M-H]", "[M+H], [M+Na], [M-H]", "[M+H]", "[M+NH4], [M+Na]"))

knitr::kable(class_adduct)
    ```
    - **Label**: "light" for natural lipids and "heavy" for U^13^C labeled lipids
    - **IMS_value**: Measured mobility (e.g. 1/K~0~ for TIMS)
    - **CCS**: Measured CCS value

3. **Import .csv file of U^13^C labeled lipid CCS library**: This needs to be a .csv file containing the ^DT^CCS~N2~ library. It is provided with the code and called "U13C_DT_CCS_library.csv".

**Code to run MobiLipid:**

```{r, include=TRUE, eval=FALSE}
# Choose R markdown which should be processed
Rmd <- tcltk::tk_choose.files(caption = "Select R markdown (.Rmd file) which should be used for data processing:",
                              filters = matrix(c("Rmd files", "Rmd"), 1, 2),
                              multi = FALSE)
Rmd_name <- basename(Rmd)
Rmd_name <- gsub("\\.Rmd$", "", Rmd_name)

# Get file path where the markdown is saved to save results in the same folder
folder_path <- dirname(Rmd)

# Import path of data which should be processed
data_import <- tcltk::tk_choose.files(caption = "Data import (measured data as CSV file):", 
                                      filters = matrix(c("csv files", "csv"), 1, 2),
                                      multi = FALSE)
filename <- basename(data_import)
filename <- gsub("\\.csv$", "", filename)

# Run R markdown
rmarkdown::render(
  paste0(Rmd),
  output_format = c("html_document", "pdf_document"),
  output_file = c(paste0(Rmd_name,"_",filename,".html"), paste0(Rmd_name,"_",filename,".pdf")))
```
